Solar wind prediction for the Solar Probe Plus orbit

### 2017-04-24 paperMVVB/sections/content.tex
finished fig 10; adjusted fig 9; upgraded fig 5

Again, the velocity gets a special treatment with the double lognormal distribution (\ref{eq:double_lognormal_fit_function}). It is known that slow and fast solar wind stream occurrence rates follow the solar cycle, but their magnitudes stay fairly stable (cite?). This time we keep the two velocity components' positions constant and vary instead their balance with the SSN:
\begin{align}
	c(ssn) &= c_a \cdot ssn + c_b\,.
\end{align}
The fit result (see Table~\ref{tab:ssn_fit_parameters}) is a model in which three years after solar cycle minimum (SSN of zero) the slow solar wind has a share of almost two-thirds and decreases further with increasing SSN (see Fig.~\ref{fig:Vdbl_SSN_ratio_b_plot}).\\
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/Vdbl_SSN_ratio_b_plot.pdf}}
	\caption{Yearly ratio between slow to fast solar wind (threshold \SI{400}{\km\per\s} over SSN (+). Fit model ratio depending on the by 3~years lagged SSN (black line). The relation results from the velocity's double lognormal fit.}
	\label{fig:Vdbl_SSN_ratio_b_plot}
\end{figure}
% 3 years after minimum (SSN of zero) the slow wind has its maximal share of 0.638(32). 3 years after maximum (SSN of 200) it has only a share of 0.278(37).\\
...compare with data\\
=> there is no specific velocity threshold between slow and fast solar wind types, the velocity ranges of both types overlap.\\

@@

With $c = 0.557$ the velocity balancing parameter is of an expected value similar to that obtained from the Helios time period (the mean SSN during the Helios period was 59, this corresponds to a $c(59) = 0.53$, see Fig.~\ref{fig:Vdbl_SSN_ratio_b_plot}).\\

@@

Implementing the orbital distance data and predicted SSN for the mission time we get SPP's estimated solar wind environment.\\
A zoom into the first and the nearest perihelia shows... (Figs.~\ref{fig:SPP_perihelia_prediction_c_plot} and \ref{fig:SPP_perihelia_prediction_nearest_c_plot}).\\
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_perihelia_prediction_c_plot.pdf}}
	\caption{Estimated solar wind parameter medians (black) and their error bands (grey) during SPP's first perihelion in 2018. For the velocity slow and fast part are plotted.}
	\label{fig:SPP_perihelia_prediction_c_plot}
\end{figure}
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_perihelia_prediction_nearest_c_plot.pdf}}
	\caption{Estimated solar wind parameter medians (black) and their error bands (grey) during SPP's nearest perihelion in 2024. For the velocity slow and fast part are plotted.}
	\label{fig:SPP_perihelia_prediction_nearest_c_plot}
\end{figure}
\subsection{Near-Sun extrapolation behavior/effects}
The extrapolation distance is only about one third of the model range, but as the parameters follow exponential change, one has to look at the logarithmic distance which is indeed one and a half times the model range.\\

>>>>

Again, the velocity gets a special treatment with the double lognormal distribution (\ref{eq:double_lognormal_fit_function}). It is known that slow and fast solar wind stream occurrence rates follow the solar cycle, but their magnitudes stay fairly stable (cite?). This time we keep the two velocity components' positions constant and vary instead their balance with the SSN:
\begin{align}
	c(ssn) &= c_a \cdot ssn + c_b\,.
\end{align}
The fit result (see Table~\ref{tab:ssn_fit_parameters}) is a model in which three years after solar cycle minimum (SSN of zero) the slow solar wind has a share of almost two-thirds and decreases further with increasing SSN (see Fig.~\ref{fig:Vdbl_SSN_ratio_d_plot}).\\
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/Vdbl_SSN_ratio_d_plot.pdf}}
	\caption{Balance of slow to fast solar wind over the by 3~years lagged SSN. The yearly ratios (+) and their weighted linear fit (solid) are obtained from OMNI data with a threshold of \SI{400}{\km\per\s}. The error bars denote the SSN standard deviation and the relative weight from the yearly data coverage. The model's balance is obtained without a threshold (dashed).}
	\label{fig:Vdbl_SSN_ratio_d_plot}
\end{figure}
% 3 years after minimum (SSN of zero) the slow wind has its maximal share of 0.638(32). 3 years after maximum (SSN of 200) it has only a share of 0.278(37).\\
The ratio of slow to fast wind obtained from yearly OMNI data via simply applying the commonly used constant velocity threshold of \SI{400}{\km\per\s} (cite?) has a flatter SSN dependence than the model's balance. Why?\\
Both should not necessary be identical, because there is no specific velocity threshold between slow and fast solar wind types, as the velocity ranges of both types overlap (cite?).\\

@@

With $c = 0.557$ the velocity balancing parameter is of an expected value similar to that obtained from the Helios time period (the mean SSN during the Helios period was 59, this corresponds to a $c(59) = 0.53$, see Fig.~\ref{fig:Vdbl_SSN_ratio_d_plot}).\\

@@

Implementing the orbital distance data and predicted SSN for the mission time we get SPP's estimated solar wind environment.\\
A zoom into the first and the nearest perihelia shows... (Figs.~\ref{fig:SPP_perihelia_prediction_c_plot} and \ref{fig:SPP_perihelia_prediction_nearest_c_plot}).\\
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_perihelia_prediction_c_plot.pdf}}
	\caption{Estimated solar wind parameter medians (black) and their error bands (grey) during SPP's first perihelion in 2018. For the velocity the SSN independent slow and fast parts are plotted.}
	\label{fig:SPP_perihelia_prediction_c_plot}
\end{figure}
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_perihelia_prediction_nearest_c_plot.pdf}}
	\caption{Estimated solar wind parameter medians (black) and their error bands (grey) during SPP's nearest perihelion in 2024. For the velocity the SSN independent slow and fast parts are plotted.}
	\label{fig:SPP_perihelia_prediction_nearest_c_plot}
\end{figure}
\subsection{Near-Sun extrapolation behavior/effects}
The extrapolation distance is only about one third of the model range, but as the parameters follow exponential change, one has to look at the logarithmic distance which is indeed one and a half times the model range.\\
